Informing policy via dynamic models: Eliminating cholera in Haiti

### 2022-10-10 ms.Rnw
Updated R code to spatPomp fitting

<<FitModel3, echo=FALSE, include=FALSE, message=FALSE>>=
set.seed(636813)

if (RUN_LEVEL == 1){

  # Global search, All parameters
  SEARCH1 <- list(NP = 20, NMIF = 3, NREPS = 3, NREPS_EVAL = 3, NP_EVAL = 25)

  # Local search, all parameters
  SEARCH2 <- list(
    TOP_N = 1, NP = 20, NMIF = 3, NREPS = 3, NREPS_EVAL = 3, NP_EVAL = 25
  )

  # Local search, unit specific parameters
  SEARCH3 <- list(TOP_N = 1, NP = 20, NMIF = 3, NREPS = 3, NREPS_EVAL = 3, NP_EVAL = 25)

  # Number of searches
  N_SEARCHES <- 3L

} else if (RUN_LEVEL == 2) {

  # Global search, all parameters
  SEARCH1 <- list(
    NP = 200, NMIF = 10, NREPS = 36, NREPS_EVAL = 36, NP_EVAL = 500
  )

  # Local search, all parameters
  SEARCH2 <- list(
    TOP_N = 2, NP = 200, NMIF = 20, NREPS = 18, NREPS_EVAL = 36, NP_EVAL = 500
  )

  # Local search, unit specific parameters
  SEARCH3 <- list(
    TOP_N = 2, NP = 200, NMIF = 10, NREPS = 18, NREPS_EVAL = 36, NP_EVAL = 500
  )

  # Number of searches
  N_SEARCHES <- 3L
} else if (RUN_LEVEL == 3) {

  # Global search, all parameters

  SEARCH1 = list(
    NBPF = 100,
    NP = 1500,
    SPAT_REGRESSION = 0.5,
    NREPS = 144,
    NP_EVAL = 2000,
    NREPS_EVAL = 36
  )

  SEARCH2 = list(
    TOP_N = 12,
    NBPF = 100,
    NP = 2000,
    SPAT_REGRESSION = 0.5,
    NREPS = 6,
    NP_EVAL = 3000,
    NREPS_EVAL = 36
  )

  SEARCH3 = list(
    TOP_N = 9,
    NBPF = 100,
    NP = 2500,
    SPAT_REGRESSION = 0.5,
    NREPS = 8,
    NP_EVAL = 5000,
    NREPS_EVAL = 36
  )

  # Total number of searches
  N_SEARCHES <- 3L
} else {
  stop("Run level must be 1, 2 or 3.")
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/haiti3_fit.rds")) {
  haiti3_fit <- readRDS("model3/run_level_3/haiti3_fit.rds")
} else {
  haiti3_fit <- bake(
    file = paste0("model3/", rl_dir, "haiti3_fit.rds"), {
      fit_haiti3(
        search1 = SEARCH1,
        search2 = SEARCH2,
        search3 = SEARCH3,
        n_searches = N_SEARCHES,
        ncores = cores
      )
    }
  )
}

rm(SEARCH1, SEARCH2, SEARCH3, N_SEARCHES)
gc()
@

@@

# MIF_mod3 <- local_MIF2_search
# rm(mif_logLik, local_MIF2_search, tLocal)

@@

# best_m3 <- ll_mod3 %>%
#   arrange(-logLik) %>%
#   slice_head(n = 1) %>%
#   pull(which)

best_m3 <- haiti3_fit[[length(haiti3_fit)]]$logLiks %>%
  filter(logLik == max(logLik, na.rm = TRUE)) %>%
  pull(which)

p3 <- haiti3_fit[[length(haiti3_fit)]]$params[best_m3, ]
gc()

@@

  h2_epi <- haiti2(region = 'before', cutoff = 10000, measure = 'log')

  coef(h2_epi) <- h2_fit$h2_params

  h3 <- haiti3_panel(start_time = "2010-10-23", B0 = TRUE)
  coef(h3) <- p3
  pp3 <- pparams(h3)
  p3u <- pp3$specific
  p3s <- pp3$shared
})
@

@@

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/SpatPompSims.rds")) {
  h3Spat_quants <- readRDS("model3/run_level_3/SpatPompSims.rds")
} else {
  h3Spat_quants <- bake(
    file = paste0('model3/', rl_dir, 'SpatPompSims.rds'), {
      h3Spat <- haiti3_spatPomp()
      coef(h3Spat) <- haitipkg:::panelParms_toSpatParms3(p3)

@@

h3Spat <- haiti3_spatPomp()
coef(h3Spat) <- haitipkg:::panelParms_toSpatParms3(p3)

>>>>

<<FitModel3, echo=FALSE, include=FALSE, message=FALSE>>=
set.seed(1859687)

# Create vectors for the unit and shared parameters
unit_specific_names <- c("betaB", "foi_add", "Binit")
shared_param_names <- c(
  "mu_B", "XthetaA", "thetaI", "lambdaR", "r", "std_W",
  "epsilon", "k", "sigma"
)
est_param_names <- c(
  unit_specific_names, shared_param_names
)

# Add unit numbers to each parameter
est_param_names_expanded <- paste0(rep(est_param_names, each = 10), 1:10)

setrw1 <- function(x) {
  if (grepl('^r[[:digit:]]+$', x)) {
    return(0.01 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.02)))
  } else {
    return(0.01)
  }
}

setrw2 <- function(x) {
  if (grepl('^r[[:digit:]]+$', x)) {
    return(0.0025 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.005)))
  } else {
    return(0.0025)
  }
}

setrw3 <- function(x) {
  if (grepl('^r[[:digit:]]+$', x)) {
    return(0.001 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.002)))
  } else {
    return(0.001)
  }
}

# First Search RW
reg_rw_1.sd <- lapply(est_param_names_expanded, setrw1)
names(reg_rw_1.sd) <- est_param_names_expanded
chol_rw_1.sd <- do.call(rw.sd, reg_rw_1.sd)

# Second Search RW
reg_rw_2.sd <- lapply(est_param_names_expanded, setrw2)
names(reg_rw_2.sd) <- est_param_names_expanded
chol_rw_2.sd <- do.call(rw.sd, reg_rw_2.sd)

# Third Search RW
reg_rw_3.sd <- lapply(est_param_names_expanded, setrw3)
names(reg_rw_3.sd) <- est_param_names_expanded
chol_rw_3.sd <- do.call(rw.sd, reg_rw_3.sd)

# Search hyper-parameters for each run-level and search
if (RUN_LEVEL == 1){
  SEARCH1 = list(
    NBPF = 5,
    NP = 50,
    SPAT_REGRESSION = 0.5,
    NREPS = 6,
    NP_EVAL = 100,
    NREPS_EVAL = 6,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 2,
    NBPF = 5,
    NP = 50,
    SPAT_REGRESSION = 0.5,
    NREPS = 3,
    NP_EVAL = 100,
    NREPS_EVAL = 6,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 2,
    NBPF = 5,
    NP = 100,
    SPAT_REGRESSION = 0.5,
    NREPS = 3,
    NP_EVAL = 100,
    NREPS_EVAL = 6,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 2L
} else if (RUN_LEVEL == 2) {
  SEARCH1 = list(
    NBPF = 50,
    NP = 1000,
    SPAT_REGRESSION = 0.5,
    NREPS = 36,
    NP_EVAL = 1000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.36,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 6,
    NBPF = 50,
    NP = 1000,
    SPAT_REGRESSION = 0.5,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 50,
    NP = 1000,
    SPAT_REGRESSION = 0.5,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 3L
} else if (RUN_LEVEL == 3) {

  SEARCH1 = list(
    NBPF = 100,
    NP = 10000,
    SPAT_REGRESSION = 0.5,
    NREPS = 72,
    NP_EVAL = 10000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 9,
    NBPF = 100,
    NP = 10000,
    SPAT_REGRESSION = 0.5,
    NREPS = 4,
    NP_EVAL = 10000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 100,
    NP = 10000,
    SPAT_REGRESSION = 0.5,
    NREPS = 6,
    NP_EVAL = 10000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 3L
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/haiti3_coupled_fit.rds")) {
  haiti3_fit <- readRDS("model3/run_level_3/haiti3_coupled_fit.rds")
} else {
  haiti3_fit <- bake(
    file = paste0("model3/", rl_dir, "haiti3_coupled_fit.rds"), {
      fit_coupled_haiti3(
        nsearches = N_SEARCHES,
        search1 = SEARCH1,
        search2 = SEARCH2,
        search3 = SEARCH3,
        ncores = cores
      )
    }
  )
}

rm(
  SEARCH1, SEARCH2, SEARCH3, N_SEARCHES, unit_specific_names, 
  shared_param_names, est_param_names, est_param_names_expanded, 
  setrw1, setrw2, setrw3, reg_rw_1.sd, reg_rw_2.sd, reg_rw_3.sd, 
  chol_rw_1.sd, chol_rw_2.sd, chol_rw_3.sd
)
gc()

@@

@@

best_m3 <- which.max(haiti3_fit[[length(haiti3_fit)]]$logLiks$logLik)

p3 <- haiti3_fit[[length(haiti3_fit)]]$params[best_m3, ]

p3_specific <- c("foi_add", "betaB", "Binit")
p3_shared <- c(
  "mu_B", "XthetaA", "thetaI", "lambdaR", "r", "std_W",
  "epsilon", "k", "sigma"
)

p3u <- matrix(nrow = length(p3_specific), ncol = 10)
rownames(p3u) <- p3_specific
p3s <- numeric(length(p3_shared))
names(p3s) <- p3_shared

for (p in p3_shared) {
  pat <- paste0("^", p, "[[:digit:]]+$")
  p3s[p] <-  mean(p3[grepl(pat, names(p3))])
}

for (p in p3_specific) {
  pat <- paste0("^", p, "[[:digit:]]+$")
  p3u[p, ] <- p3[grepl(pat, names(p3))]
}

rm(p3_specific, p3_shared, p, pat)

@@

  h2_epi <- haiti2(region = 'before', cutoff = 10000, measure = 'log')
  coef(h2_epi) <- h2_fit$h2_params

  h3Spat <- haiti3_spatPomp()
  coef(h3Spat) <- p3
})
@

@@

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/SpatPompSims.rds")) {
  h3Spat_quants <- readRDS("model3/run_level_3/SpatPompSims.rds")
} else {
  h3Spat_quants <- bake(
    file = paste0('model3/', rl_dir, 'SpatPompSims.rds'), {
      # h3Spat <- haiti3_spatPomp()
      # coef(h3Spat) <- haitipkg:::panelParms_toSpatParms3(p3)

@@

# h3Spat <- haiti3_spatPomp()
# coef(h3Spat) <- haitipkg:::panelParms_toSpatParms3(p3)
